---
title: FDA-Study-1 comparison between semi-automatic annotations and automatic Tpeak
  and Tend delineation method
author: "Jose Vicente <Jose.VicenteRuiz@fda.hhs.gov>"
date: "May, 2017"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, 
                      echo=FALSE, warning=FALSE, message=FALSE,error=F)
```

**Disclaimer**

This code does not necessarily reflect any position of the Government or the Food and Drug Administration.

This software and documentation (the "Software") were developed at the Food and Drug Administration (FDA) by employees of the Federal Government in the course of their official duties. Pursuant to Title 17, Section 105 of the United States Code, this work is not subject to copyright protection and is in the public domain. Permission is hereby granted, free of charge, to any person obtaining a copy of the Software, to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, or sell copies of the Software or derivatives, and to permit persons to whom the Software is furnished to do so. FDA assumes no responsibility whatsoever for use by other parties of the Software, its source code, documentation or compiled executables, and makes no guarantees, expressed or implied, about its quality, reliability, or any other characteristic. Further, use of this code in no way implies endorsement by the FDA or confers any advantage in regulatory decisions. Although this software can be redistributed and/or modified freely, we ask that any derivative works bear some notice that they are derived from it, and any modified versions bear some notice that they have been modified.

```{r loadpackages}
library(tidyverse);
library(lme4);
library(lsmeans);
library(pander);
library(BlandAltmanLeh);
library(ggExtra);
library(gridExtra);

```

```{r loaddata,results='asis'}
if(!exists('filterwasenabled')){
  filterwasenabled <- 1
}
if(filterwasenabled==0){
  pandoc.header('T wave delineated *as is* (no filter applied)',level=1)
} else {
  pandoc.header('T wave delineated after applying filter as described in documentation',level=1)
}
cat('\n')
cat('* * * * \n')

if(!exists('theref')){
  theref <- 'reference'
}
if(!exists('themethod')){
  themethod <- 'algorithm'
}


if(!file.exists('SCR-002.Clinical.Data.csv')){
    res <- tryCatch(download.file('https://www.physionet.org/physiobank/database/ecgrdvq/SCR-002.Clinical.Data.csv',
                              destfile="./SCR-002.Clinical.Data.csv",
                              method="auto"),
                error=function(e) 1)
}
studydata <- read.csv('SCR-002.Clinical.Data.csv', as.is = T, stringsAsFactors = F) %>% select(-ERD_30,-LRD_30,- Twave_amplitude,-Twave_asymmetry,-Twave_flatness);
referenceannotations <- read.csv('dbannotations.csv', as.is = T, stringsAsFactors = F);
referenceannotations <- referenceannotations %>% mutate(annotator='reference') %>% mutate(PON=ifelse(PON<0,NA,PON),QON=ifelse(QON<0,NA,QON),QOFF=ifelse(QOFF<0,NA,QOFF),TPEAK=ifelse(TPEAK<0,NA,TPEAK),TPPEAK=ifelse(TPPEAK<0,NA,TPPEAK),TOFF=ifelse(TOFF<0,NA,TOFF),ERROR='') %>% select(-PON);
algorithmannotations <- read.csv('results.csv', as.is = T, stringsAsFactors = F);
algorithmannotations <- algorithmannotations %>% filter(FILTER==filterwasenabled) %>% mutate(annotator='algorithm')  %>% mutate(TPPEAK=ifelse(TPPEAK<0,NA,TPPEAK)) %>% group_by(RECORD) %>% mutate(EGREFID=strsplit(RECORD,'/')[[1]][4]) %>% ungroup() %>% data.frame() %>% rename(TOFF=TEND) %>% select(-FILTER,-RR)
annotationsdata <- bind_rows(referenceannotations,algorithmannotations) %>% filter(ERROR=='') %>% select(-ERROR,-RPEAK) 
intervalsfromannotations <- annotationsdata %>%  mutate(QT=TOFF-QON,QRS=QOFF-QON,JTPEAK=TPEAK-QOFF,TPEAKTEND=TOFF-TPEAK,TPEAKTPEAKP=TPPEAK-TPEAK) %>% ungroup() %>% select(EGREFID,annotator,QT,QRS,JTPEAK,TPEAKTEND,TPEAKTPEAKP)
annotationsdatalong <- annotationsdata %>% gather(Annotation.Type,Value.ms,c(QON,QOFF,TPEAK,TPPEAK,TOFF))
intervalsfromstudydata <- studydata %>% mutate(annotator='referencecsv') %>% select(EGREFID,annotator,QT,QRS,JTPEAK,TPEAKTEND,TPEAKTPEAKP)

allintervals <- bind_rows(intervalsfromstudydata,intervalsfromannotations)
```

# Tpeak, Tpeak' and Tend annotations differences

```{r annotationdifferences, results="asis"}
annotationsdatawide <- annotationsdatalong %>% spread(annotator,Value.ms) %>% rename(param=Annotation.Type) %>% filter(is.element(param,c('TPEAK','TPPEAK','TOFF')))
#Compute BlandAltman stats and plots
bastats <- list()
baplots <- list()
params <- unique(annotationsdatawide$param)
for(p in params){
  tmp <- annotationsdatawide %>% filter(param==p)
  bastats$annots[[p]] <- bland.altman.stats(tmp[[themethod]],tmp[[theref]])
  baplots$annots[[p]] <- ggMarginal((bland.altman.plot(tmp[[themethod]],tmp[[theref]],graph.sys = "ggplot2") + theme_bw() + ggtitle(paste0(p,"\n",theref," vs. ",themethod))), type = "histogram", size=4,binwidth=5)
  pandoc.header(p,level=2)
  print(baplots$annots[[p]])
  cat('\n')
  cat('\n')
  pandoc.table(bastats$annots[[p]]$lines)
  cat('\n')
  cat('\n')
}
```

# ECG intervals differences

```{r assessdiferences}
#Transpose data.frame from wide to long
allintervalslong <- allintervals %>% gather(param,value,3:7)
allintervalsforcheck <- allintervalslong %>% spread(annotator,value) %>% mutate(referencecsv_reference=referencecsv-reference,referencecsv_algorithm=referencecsv-algorithm,reference_algorithm=reference-algorithm)

#Exclude PR interval because P onset was not included in the output of twave delineator
allintervalsforcheck <- allintervalsforcheck %>% filter(param!='PR')

#Compute BlandAltman stats and plots
#bastats <- list()
#baplots <- list()
params <- unique(allintervalsforcheck$param)
for(p in params){
  tmp <- allintervalsforcheck %>% filter(param==p)
  bastats$referencecsv_reference[[p]] <- bland.altman.stats(tmp$referencecsv,tmp[[theref]])
  baplots$referencecsv_reference[[p]] <- ggMarginal((bland.altman.plot(tmp$referencecsv,tmp[[theref]],graph.sys = "ggplot2") + theme_bw() + ggtitle(paste0(p,"\n referenceCSV vs. ",theref))), type = "histogram", size=4,binwidth=5)
  bastats$reference_algorithm[[p]] <- bland.altman.stats(tmp[[theref]],tmp[[themethod]])
  baplots$reference_algorithm[[p]] <- ggMarginal((bland.altman.plot(tmp[[theref]],tmp[[themethod]],graph.sys = "ggplot2") + theme_bw() + ggtitle(paste0(p,"\n",theref," vs. ",themethod))), type = "histogram", size=4,binwidth=5)
}
```

## Intervals from referencecsv vs. reference annotations

This section is provided just for quality control purposes.

```{r plotdifferences, results="asis"}
for(p in params){
  pandoc.header(p,level=3)
  print(baplots$referencecsv_reference[[p]])
  cat('\n')
  cat('\n')
  pandoc.table(bastats$referencecsv_reference[[p]]$lines)
  cat('\n')
  cat('\n')
}
```

## Intervals from reference vs. algorithm annotations

```{r plotalgorithmdifferences, results='asis'}
for(p in params){
  pandoc.header(p,level=3)
  print(baplots$reference_algorithm[[p]])
  cat('\n')
  cat('\n')
  pandoc.table(bastats$reference_algorithm[[p]]$lines)
  cat('\n')
  cat('\n')
}
```

# Baseline measures

```{r baselineandsingeldeltas, results="asis"}
timeinfo <- studydata %>% select(EGREFID, RANDID, EXTRT, BASELINE, TPT, RR)

allintervalswithtime <- merge(timeinfo,allintervals,by=c('EGREFID')) %>% arrange(RANDID,annotator,EXTRT,TPT) %>% gather(param,value,8:12) %>% filter(is.element(param,c('QT','JTPEAK','TPEAKTEND')))

#Heart rate correction
allintervalswithtimec <- allintervalswithtime %>% mutate(value=ifelse(param=='QT',(value/((RR/1000)^(1/3))), ifelse(param=='JTPEAK',(value/((RR/1000)^(0.58))),value))) %>% mutate(param=ifelse(param=='QT','QTc', ifelse(param=='JTPEAK','JTPEAKc',param)))

#Baseline
bldata <- allintervalswithtimec %>% filter(BASELINE=='Y') 
blsummarytable <- bldata %>% group_by(annotator,RANDID,param) %>% summarise(SUBJAVG=mean(value,na.rm=T)) %>% ungroup() %>% group_by(annotator,param) %>% summarise(AVG=mean(SUBJAVG,na.rm=T),SD=sd(SUBJAVG,na.rm=T),NSubjects=n()) %>% mutate(txt=sprintf("%.1f \u00b1 %.1f",round(AVG,1),round(SD,1)))
cat('\n')
cat('\n')
#Exclude reference intervals from csv file
pandoc.table(blsummarytable %>% filter(annotator!='referencecsv') %>% select(annotator,param,txt) %>% spread(annotator,txt))
cat('\n')
cat('\n')

#single deltas
#Compute changes from baseline (single deltas) for each timepoint per subject and biomarker (differences of the mean values of the triplicate ECG extractions)
avgstudydatalong <- allintervalswithtimec %>% group_by(annotator,RANDID, EXTRT, TPT, param) %>% mutate(avgval=mean(value,na.rm=T)) %>% select(-value,-EGREFID) %>% rename(value=avgval) %>% ungroup() %>% select(-RR) %>% distinct()
bldatameasures <- avgstudydatalong %>% filter(BASELINE=='Y') %>% select(annotator,RANDID, EXTRT, param, value) %>% rename(BL=value)
singledeltadatalong <- merge(avgstudydatalong,bldatameasures) %>% filter(BASELINE!='Y') %>% mutate(BLCHG=value-BL)

```

# Single deltas BA plots (similar to Figure 3 in the paper)

```{r basingledeltas, results='asis'}
params <- unique(singledeltadatalong$param)
for(p in params){
  tmp <- singledeltadatalong %>% filter(param==p) %>% select(-value,-BL) %>% spread(annotator,BLCHG)
  bastats$sd_reference_algorithm[[p]] <- bland.altman.stats(tmp[[theref]],tmp[[themethod]])
  baplots$sd_reference_algorithm[[p]] <- ggMarginal((bland.altman.plot(tmp[[theref]],tmp[[themethod]],graph.sys = "ggplot2") + theme_bw() + ggtitle(paste0(p,"\n",theref," vs. ",themethod))), type = "histogram", size=4,binwidth=5)
  pandoc.header(p,level=2)
  print(baplots$sd_reference_algorithm[[p]])
  cat('\n')
  cat('\n')
  pandoc.table(bastats$sd_reference_algorithm[[p]]$lines)
  cat('\n')
  cat('\n')
}
```

## PLOS ONE paper Figure 3

**Note:** Black horizontal dashed lines in this plot show 95% confidence intervals of the differences while red dashed horizontal lines show the standard deviation of the differences similar to Figure 3 in Johannesen et al PLoS One 2016 [https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5201230/](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5201230/).

```{r plosonefigure,fig.width=12,fig.height=4}
params <- unique(singledeltadatalong$param)
#ponexlims <- list("JTPEAKc"=c(-90,90),"TPEAKTEND"=c(-20,150),"QTc"=c(-40,140))
ponestring <- list()
pmean <- list()
psd <- list()
for(p in params){
  tmp <- singledeltadatalong %>% filter(param==p) %>% select(-value,-BL) %>% spread(annotator,BLCHG)
  bastats$sd_reference_algorithm[[p]] <- bland.altman.stats(tmp[[theref]],tmp[[themethod]])
  pmean[[p]] <- round(mean(bastats$sd_reference_algorithm[[p]]$diffs),1)
  psd[[p]] <- round(sd(bastats$sd_reference_algorithm[[p]]$diffs),1)
  ponestring[[p]] <- paste0(pmean[[p]], ' (',psd[[p]],')')
  baplots$sd_reference_algorithm[[p]] <- ggMarginal((bland.altman.plot(tmp[[theref]],tmp[[themethod]],graph.sys = "ggplot2") + theme_bw() + ggtitle(paste0(p,"\n",theref," vs. ",themethod)) + scale_y_continuous(limits=c(-150,150),oob=scales::rescale_none) +
                                                       #scale_x_continuous(limits=ponexlims[[p]],oob=scales::rescale_none) +
                                                       annotate("text",x=20,y=-75,label=ponestring[[p]]) + 
                                                       geom_hline(yintercept = c(-psd[[p]],psd[[p]]),linetype=2,colour='red')), type = "histogram", size=4,binwidth=5)
}
grid.arrange(baplots$sd_reference_algorithm$JTPEAKc,
             baplots$sd_reference_algorithm$TPEAKTEND,
             baplots$sd_reference_algorithm$QTc,
             nrow=1)
figurename <- paste0('plosone-',ifelse(filterwasenabled,"Filtered","No-Filtered"),'.png')
p <- png(figurename,width=600,height=200)
grid.arrange(baplots$sd_reference_algorithm$JTPEAKc,
             baplots$sd_reference_algorithm$TPEAKTEND,
             baplots$sd_reference_algorithm$QTc,
             nrow=1)
p<-dev.off()
```

# R session information

```{r techappendix}
sessionInfo()
```