---
title: "ISCE J-Tpeak initiative template report example"
author: "Jose Vicente <Jose.VicenteRuiz@fda.hhs.gov>"
date: "February 2016 (minor updates on May 2017)"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE,error=FALSE)
```


**Disclaimer**

This code does not necessarily reflect any position of the Government or the Food and Drug Administration.

This software and documentation (the "Software") were developed at the Food and Drug Administration (FDA) by employees of the Federal Government in the course of their official duties. Pursuant to Title 17, Section 105 of the United States Code, this work is not subject to copyright protection and is in the public domain. Permission is hereby granted, free of charge, to any person obtaining a copy of the Software, to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, or sell copies of the Software or derivatives, and to permit persons to whom the Software is furnished to do so. FDA assumes no responsibility whatsoever for use by other parties of the Software, its source code, documentation or compiled executables, and makes no guarantees, expressed or implied, about its quality, reliability, or any other characteristic. Further, use of this code in no way implies endorsement by the FDA or confers any advantage in regulatory decisions. Although this software can be redistributed and/or modified freely, we ask that any derivative works bear some notice that they are derived from it, and any modified versions bear some notice that they have been modified.

#Abstract
This script was written for ISCE 2017 J-Tpeak initiative and it is based on previous analysis published in Vicente et al JAHA 2015 (https://doi.org/10.1161/JAHA.114.001615). This script generates time-profile and exposure-response plots for J-Tpeak measures assessed on ECGs from FDA's prospective clinical study NCT01873950. Clinical data and ECGs of this study are available at http://thew-project.org/Database/E-OTH-12-5232-020.html and https://physionet.org/physiobank/database/ecgrdvq/.

#Load libraries and declare utility functions

```{r packageload, message=FALSE, warning=FALSE}
############################################################################
#Load required libraries
############################################################################

## Graphics and data wrangling
library(tidyverse);
library(pander);
library(ggplot2);
library(gridExtra);

## Model fitting
library(lme4);

## Estimation of model parameter p-values and CI
library(pbkrtest);

## Estimation of double deltas (e.g. ddJTpeakc) by contrasts
library(lsmeans);

## Summary function for using dplyr::summarize_(). 
my.sum.fun<-funs(mean   = mean(. ,na.rm=T),
                 median = median(. , na.rm=T),
                 min    = min(. ,na.rm=T),
                 max    = max(. ,na.rm=T),
                 sd     = sd(. ,na.rm=T), 
                 n      = sum(!is.na(.)),
                 se     = sd(. ,na.rm=T)/sqrt(sum(!is.na(.))),
                 LCL    = mean(. ,na.rm=T)+qnorm(0.05)*(sd(. ,na.rm=T)/sqrt(sum(!is.na(.)))),
                 UCL    = mean(. ,na.rm=T)+qnorm(0.95)*(sd(. ,na.rm=T)/sqrt(sum(!is.na(.)))))

## Color palette for plots
treatmentcolors <- c("Dofetilide"="#999999","Quinidine"="#E69F00","Ranolazine"= "#56B4E9","Verapamil"="#009E73", "Placebo"="#000000");
```

#Load and prepare the data for analysis

## Load data from Physionet

Load data set from https://physionet.org/physiobank/database/ecgrdvq/SCR-002.Clinical.Data.csv and prepare data.frame variables for analysis.

```{r loaddata}
## Load dataset. This line can be edited to load a local copy of the csv file instead. (e.g.: use 'dataset/SCR-002.Clinical.Data.csv' in read.csv)
study1data <- read.csv('https://physionet.org/physiobank/database/ecgrdvq/SCR-002.Clinical.Data.csv', as.is = T, stringsAsFactors = F);

## Subset clinical data
clinicaldata <- study1data %>% select(RANDID,SEX,AGE,VISIT,ARMCD,EXTRT,TPT,BASELINE,PCTEST,PCSTRESN,PCSTRESU,EGREFID) %>% rename(ECGID=EGREFID)

## Scale concentrations as appropriate to avoid potential model convergence issues with lmer later
clinicaldata <- clinicaldata %>% mutate(PCSTRESN=ifelse(PCTEST!='Verapamil',PCSTRESN/1000,PCSTRESN), PCSTRESU=ifelse(PCTEST=='Dofetilide','ng/mL',ifelse(PCTEST!='Verapamil','ug/mL',PCSTRESU)))

## Populate Placebo PK info with 0 concentration
clinicaldata <- clinicaldata %>% mutate(PCTEST=ifelse(EXTRT=='Placebo','Placebo',PCTEST), PCSTRESN=ifelse(EXTRT=='Placebo',0,PCSTRESN), PCSTRESU=ifelse(EXTRT=='Placebo','n/a',PCSTRESU))

## Subset ECG data
physionetdata <- study1data %>% rename(ECGID=EGREFID,JT_VM=JTPEAK) %>% select(ECGID,RR,JT_VM) %>% mutate(JT_II=NA,JT_V5=NA,	BS_FLAG=2,PARTICIPANT='FDA') %>% select(PARTICIPANT,BS_FLAG,ECGID,RR,JT_II,JT_V5,JT_VM) %>% arrange(PARTICIPANT,ECGID)
```

## Load data from submissions

Load additional datasets from submissions to the J-Tpeak initiative.

```{r loadsubmissions}
## Populate list of received files
receviedfiles <- c('submissions/filtered.csv','submissions/nofiltered.csv') # Initialize received files to load additional datasets submitted by participants to the initiative (e.g. with c('submissions/participant1.csv','submissions/participant2.csv'))
allsubmissionsdata <- NULL
## Load and merge the data into one data.frame
for(submission in receviedfiles){
  ## Load annotations from the submissions
  submissiondata <- read.csv(submission, as.is = T, stringsAsFactors = F);
  ## Merge submitted datasets
  allsubmissionsdata <- bind_rows(allsubmissionsdata,submissiondata)
}
## Merge physionet and submitted datasets
allsubmissionsdata <- bind_rows(physionetdata,allsubmissionsdata)

## Transpose to long format
allsubmissionsdatalong <- allsubmissionsdata %>% gather(param,value,5:7)
```

## Heart rate correction

Correct J-Tpeak measures for heart rate effects. In this example we use the model and coefficient published in Johannesen et al. Clin Pharm Ther 2014 (https://doi.org/10.1038/clpt.2013.238).

```{r hrcorrection}
## Apply heart rate correction
allsubmissionsdatalonghr <- allsubmissionsdatalong %>% group_by(PARTICIPANT,ECGID) %>% mutate(param=paste0(param,'c'),value=round(value/((RR/1000)^(0.58)),0)) %>% ungroup() %>% data.frame()

## Bind all rows from orignal and hear rate corrected data.frames
analysisdataset <- bind_rows(allsubmissionsdatalong,allsubmissionsdatalonghr) %>% arrange(PARTICIPANT,ECGID,param)
```

## Changes from baseline (single deltas: &Delta;)

```{r singledeltas}
## Merge clinical data with JTpeak measures
analysisdataset <- merge(clinicaldata,analysisdataset) %>% arrange(RANDID,EXTRT,PARTICIPANT,param,TPT)

## Compute average values
avganalysisdataset <- analysisdataset %>% group_by(RANDID,EXTRT,PARTICIPANT,param,TPT) %>% mutate(RR=round(mean(RR,na.rm=T),0),value=round(mean(value,na.rm=T),0)) %>% ungroup() %>% select(-ECGID) %>% distinct()

## Subset baseline rows
baselinedata <- avganalysisdataset %>% filter(BASELINE=='Y') %>% select(RANDID,EXTRT,PARTICIPANT,param,value) %>% rename(BL=value) 

## Merge BL measures with no-baseline rows in analysisdataset
singledeltasdataset <- merge(avganalysisdataset %>% filter(BASELINE!='Y'),baselinedata,by=c('PARTICIPANT','RANDID','EXTRT','param'))

## Compute change from baseline
singledeltasdataset <- singledeltasdataset %>% mutate(BLCHG=value-BL) %>% select(RANDID,SEX,AGE,VISIT,ARMCD,EXTRT,TPT,PCTEST,PCSTRESN,PCSTRESU,PARTICIPANT,BS_FLAG,RR,param,value,BL,BLCHG)

## Assign factor levels so Placebo can be used as contrast later in the lme models
singledeltasdataset$EXTRT <-  factor(singledeltasdataset$EXTRT,levels=c('Placebo','Dofetilide','Quinidine Sulph','Ranolazine','Verapamil HCL'),labels=c('Placebo','Dofetilide','Quinidine','Ranolazine','Verapamil'));
singledeltasdataset$EXTRT <- relevel(singledeltasdataset$EXTRT,ref='Placebo')

## Similarly, assign factors for timepoints so they can be managed by ggplot later
singledeltasdataset <- singledeltasdataset %>% mutate(TPTN=TPT)
singledeltasdataset <- singledeltasdataset %>% mutate(TPT=paste0('Time',TPT))
singledeltasdataset$TPT <- factor(singledeltasdataset$TPT,levels=unique(singledeltasdataset$TPT),ordered=T)

```

## Placebo-corrected changes from baseline (double deltas: &Delta;&Delta;)

```{r doubledeltas}
## Subset placebo rows
placebodata <- singledeltasdataset %>% filter(EXTRT=='Placebo') %>% rename(PBORR=RR,PBO=value,PBOBL=BL,PBOBLCHG=BLCHG) %>% select(RANDID,TPT,TPTN,PARTICIPANT,param,PBORR,PBO,PBOBL,PBOBLCHG)

## Merge PBO measures with no-placebo rows in singledeltasdataset
doubledeltasdataset <- merge(singledeltasdataset %>% filter(EXTRT!='Placebo'),placebodata,by=c('PARTICIPANT','RANDID','TPT','TPTN','param')) 

## Compute placebo corrected change from baseline (double delta)
doubledeltasdataset <- doubledeltasdataset %>% mutate(DDCHG=BLCHG-PBOBLCHG) %>% select(RANDID,SEX,AGE,VISIT,ARMCD,EXTRT,TPT,TPTN,PCTEST,PCSTRESN,PCSTRESU,PARTICIPANT,BS_FLAG,RR,PBORR,param,value,BL,BLCHG,PBO,PBOBL,PBOBLCHG,DDCHG) %>% arrange(RANDID,EXTRT,TPT,TPTN,PARTICIPANT,param)
```

# Statistical analysis

## Pharmacodynamic time profiles

```{r dataforexample}
## Subset data to be analized in this example
if(!exists('participant')){
  participant <- 'FDA' # Select participant based on PARTICIPANT column of submitted datasets
}
if(!exists('biomarker')){
  biomarker <- 'JT_VMc' # Select ECG biomarker column
}

## Initialize data.frame for single-delta analysis
example.df <- singledeltasdataset %>% filter(PARTICIPANT==participant & param==biomarker) 

## Initialize data.frame foror double-delta analysis
ddexample.df <- doubledeltasdataset %>% filter(PARTICIPANT==participant & param==biomarker) 
```

### &Delta; time profile

The code below requires a dataframe with at least the following columns:

* RANDID: Subject identifier
* EXTRT:  Treatment arm
* ARMCD:  Treatment sequence
* VISIT:  Visit number
* TPT:    Time point
* BLCHG:  Change from baseline


```{r sdtimeanalysis, results='asis'}
## Set lsm.options to use 'Kenward-Roger' method
lsm.options(lmer.df = "Kenward-Roger" )

## Fit lme model
time.fit <- lmer(BLCHG ~ VISIT + ARMCD + TPT + EXTRT + EXTRT:TPT + (1|RANDID) ,data = example.df)

## Get reference grid from the time model
time.ref.grid <- ref.grid(time.fit,data=time.fit@frame,lsm.options(pbkrtest.limit = nrow(example.df)+1))

## Generate LS-Means estimated of baseline adjusted change (e.g. &Delta;JT-peak) per TPT timepoint and EXTRT treatment
time.lsm<-lsmeans::lsmeans(time.ref.grid, c("TPT", "EXTRT"))

## Generate (single delta) predictions with CIs
time.pred<-summary(time.lsm, level=0.95)

## Merge time values back
time.pred <- merge(time.pred,example.df %>% select(TPT,TPTN) %>% distinct())

## Plot single delta changes
ggplot(time.pred,aes(TPTN,lsmean,color=EXTRT,ymin=lower.CL, ymax=upper.CL)) + geom_point() + geom_line() + geom_errorbar() + geom_hline(yintercept = 0) + theme_bw() + ggtitle('Changes from baseline') + theme(legend.position = 'bottom',legend.title=element_blank()) + ylab(expression(paste(Delta,"  \u00B1  95% CI (ms)"))) + xlab('Time postdose (hrs)') + scale_color_manual(values=treatmentcolors) + scale_fill_manual(values=treatmentcolors)

```

### &Delta;&Delta; time profile

The code below requires a dataframe with at least the following columns:

* RANDID: Subject identifier
* EXTRT:  Treatment arm
* ARMCD:  Treatment sequence
* VISIT:  Visit number
* TPT:    Time point
* DDCHG:  Placebo corrected change from baseline


```{r ddtimeanalysis2, results='asis'}
## Fit lme model
time.fit.dd <- lmer(DDCHG ~ VISIT + ARMCD + TPT + EXTRT + EXTRT:TPT + (1|RANDID) ,data = ddexample.df)

## Get reference grid from the time model
time.ref.grid.dd <- ref.grid(time.fit.dd,data=time.fit.dd@frame,lsm.options(pbkrtest.limit = nrow(ddexample.df)+1))  

## Generate LS-Means estimated of placebo corrected change from baseline (e.g. &Delta;&Delta;JT-peak) per TPT timepoint and EXTRT treatment
time.lsm.dd<-lsmeans::lsmeans(time.ref.grid.dd, c("TPT", "EXTRT"))

## Generate (delta delta) predictions with CIs
time.pred.dd<-summary(time.lsm.dd, level=0.95)

## Merge time values back
time.pred.dd <- merge(time.pred.dd,ddexample.df %>% select(TPT,TPTN) %>% distinct())

## Plot double delta changes 
ggplot(time.pred.dd,aes(TPTN,lsmean,color=EXTRT,ymin=lower.CL, ymax=upper.CL)) + geom_point() + geom_line() + geom_errorbar() + geom_hline(yintercept = 0) + theme_bw() + ggtitle('Placebo corrected changes from baseline') + theme(legend.position = 'bottom',legend.title=element_blank()) + ylab(expression(paste(Delta,Delta," \u00B1  95% CI (ms)")))+ xlab('Time postdose (hrs)') + scale_color_manual(values=treatmentcolors) + scale_fill_manual(values=treatmentcolors) 
```

## Exposure-response

The exposure-response analysis code below requires a data.frame containing at least the following columns:

* RANDID: Subject identifier
* EXTRT:  Treatment arm
* PCSTRESN: Drug concentration
* PCSTRESU: Drug concentration units
* DDCHG: Placebo corrected change from baseline


### Exposure-response analysis

```{r eranalysis,fig.width=7.5, fig.height=7.5, results='asis'}
## Fit ER models
conc.models <- NULL
conc.preds.dd <- NULL
for(treatment in unique(ddexample.df$EXTRT)){
  ## Subset treatment data
  dchgdf <- ddexample.df %>% filter(EXTRT==treatment)
  tryCatch({
    ## Fit model 
    conc.fit.tmp <- lmer(DDCHG ~ PCSTRESN + (1+PCSTRESN|RANDID) ,data = dchgdf)
    
    ## Add model fits to master conc.models data.frame
    conc.models <- bind_rows(conc.models, data.frame() %>% do(model=conc.fit.tmp) %>% mutate(EXTRT=treatment))
    
    ## Estimate changes using fitted model throughout the entire range or observed concentration (i.e. 0 to iCmax)
    ## Compute individual maximum concentration 
    iCmax <- max(dchgdf$PCSTRESN,na.rm=T)
    ## population Cmax
    pCmax <- (dchgdf %>% group_by(PCSTRESU,RANDID) %>% summarise(iCmax=max(PCSTRESN,na.rm=T)) %>% ungroup() %>% group_by(PCSTRESU) %>% summarise(Cmax=mean(iCmax)) %>% select(Cmax))$Cmax
    ## Build list of concentrations for predictions (entire range of observed concentration and pCmax)
    concs.df <- (data.frame(PCSTRESN=c(pCmax,signif(seq(from=0, to =iCmax,length.out=100),3))) %>% mutate(isCmax=PCSTRESN==pCmax) %>% arrange(PCSTRESN))
    ## Get reference grid from the concentration response model throughout the entire range of observed concentration
    conc.ref.grid.tmp <- ref.grid(conc.fit.tmp,
                                  at=list(PCSTRESN=unique(concs.df$PCSTRESN)),
                                  data=conc.fit.tmp@frame)  
    
    ## Generate LS-Means estimated of placebo corrected changes from baseline (e.g. &Delta;&Delta;JT-peak) 
    conc.lsm.tmp<-lsmeans::lsmeans(conc.ref.grid.tmp, c("PCSTRESN"))
    ## Generate &Delta;&Delta; mean predictions with CIs
    tryCatch({
      conc.pred.tmp<-summary(conc.lsm.tmp, level=0.95)
      conc.pred.tmp$EXTRT <- treatment
      conc.pred.tmp <- conc.pred.tmp %>% mutate(isCmax=PCSTRESN==pCmax)
    }, error = function(e) {
      cat(paste0('There was an error computing the predictions from the REML model for ', treatment,'.\n'))
    });
    ## Add predticted estimates and CIs to master conc.preds.dd data.frame
    conc.preds.dd <- bind_rows(conc.preds.dd,conc.pred.tmp)
  }, error = function(e) {
    cat(paste0('There was an error fitting the model and/or computing the lsmeans for ', treatment,'.\n'))
  });
}
```

### Exposure-response plots

```{r eranalysisplots,fig.width=7.5, fig.height=7.5, results='asis'}
## Compute means and 95CIs of observed double deltas grouped by concentration deciles (10 bins)
bin.ecgpk<- ddexample.df %>%
  filter(!is.na(PCSTRESN)) %>%
  filter(EXTRT!='Placebo') %>%
  group_by(EXTRT) %>%
  mutate(Decile=ntile(PCSTRESN, 10)) %>% 
  group_by(EXTRT,Decile) %>%
  summarise_each(my.sum.fun, PCSTRESN, DDCHG) %>% ungroup()

## Select only columns needed for plotting
plot.bins <- bin.ecgpk %>% 
  select(EXTRT,Decile,PCSTRESN_min,PCSTRESN_max,PCSTRESN_median) %>%
  distinct() 
## Arbitrary values for Y axis range in the plots. Note that observed and predicted values out of the yl range will not be plotted. 
yl <- c(-10,55)
## yloc: Y coordinate for horizontal bar showing concentration bins mean values and ranges
yloc <- -10;

## Compute population's Cmax for each treatment
Cmaxes <- ddexample.df %>% group_by(EXTRT,PCSTRESU,RANDID) %>% summarise(iCmax=max(PCSTRESN,na.rm=T)) %>% ungroup() %>% group_by(EXTRT,PCSTRESU) %>% summarise(Cmax=mean(iCmax)) %>% select(EXTRT,Cmax,PCSTRESU)
## Compute DDCHG at Cmax
ddchgatcmax <- conc.preds.dd %>% filter(isCmax)

## Genarate ER plots
dd.preds.plots <- list()
for(treatment in unique(ddexample.df$EXTRT)){
  ddexample.df.trta <- ddexample.df %>% filter(complete.cases(.),EXTRT==treatment)
  conc.preds.dd.trta <- conc.preds.dd %>% filter(EXTRT==treatment)
  bin.ecgpk.trta <- bin.ecgpk  %>% filter(EXTRT==treatment)
  plot.bins.trta <- plot.bins %>% filter(EXTRT==treatment)
  Cmaxes.trta <- Cmaxes %>% filter(EXTRT==treatment)
  
  dd.preds.plots[[treatment]] <- ggplot()+
    geom_point(data=ddexample.df.trta, aes(x=PCSTRESN, y=DDCHG, col=EXTRT), alpha=0.2)+ # Print observed dd values
    geom_ribbon(data=conc.preds.dd.trta, aes(x=PCSTRESN, ymin=lower.CL, ymax=upper.CL,fill=EXTRT), alpha=0.4)+ # Print model's prediction CIs
    geom_line(data=conc.preds.dd.trta,  aes(x=PCSTRESN, y=lsmean, col=EXTRT), size=1)+ # Print model's predicted mean
    #geom_pointrange(data=conc.preds.dd.trta %>% filter(isCmax),  aes(x=PCSTRESN, y=lsmean,ymin=lower.CL,ymax=upper.CL), color='black',size=0.5)+ # Print model's predicted mean for Cmax
    geom_pointrange(data=bin.ecgpk.trta , aes(x=PCSTRESN_median, ymax=DDCHG_UCL, ymin=DDCHG_LCL, y=DDCHG_mean,col=EXTRT),size=0.25)+ # Vertical bars and dots with mean and 95 CIs of observed dd values per concentration decile
    geom_point(data=plot.bins.trta, y=yloc,aes(x=PCSTRESN_max,col=EXTRT), shape="|", size=3)+ # Vertical segments on horizontal bar with concentration deciles (deciles' upper limits)
    geom_point(data=plot.bins.trta, y=yloc,aes(x=PCSTRESN_min,col=EXTRT), shape="|", size=3)+  # Vertical segments on horizontal bar with concentration deciles  (deciles' lower limits)
    geom_line(data=plot.bins.trta, y=yloc,aes(x=PCSTRESN_min,col=EXTRT), size=0.25)+ # Horizontal segments on horizontal bar with concentration deciles (deciles' lower limits)
    geom_line(data=plot.bins.trta, y=yloc,aes(x=PCSTRESN_max,col=EXTRT), size=0.25)+ # Horizontal segments on horizontal bar with concentration deciles (deciles' upper limits)
    geom_point(data=plot.bins.trta, y=yloc,aes(x=PCSTRESN_median,col=EXTRT), size=1)+ # Median value dots on horizontal bar with concentration deciles
    ylab(expression(paste(Delta,Delta," \u00B1  95% CI (ms)")))+ # Y axis label
    xlab(paste0(treatment,' (',unique(ddexample.df.trta$PCSTRESU),')'))+ # Set x axis label
    scale_y_continuous(limits=yl,oob=scales::rescale_none) + # Y axis range
    scale_color_manual(values=treatmentcolors) + scale_fill_manual(values=treatmentcolors) + # Set treatment colors
    geom_hline(data=plot.bins.trta,aes(yintercept=10), linetype=3)+ # Horizontal doted line at 10 ms
    geom_hline(data=plot.bins.trta,aes(yintercept=0), linetype=2)+ # Horizontal dashed line at 0 ms
    geom_vline(data=Cmaxes.trta,aes(xintercept=Cmax))+ # Vertical line showing treatment Cmax
    theme_bw() +
    theme(legend.position="none",legend.title=element_blank())+
    ggtitle(treatment)
}

## Print plots
grid.arrange(dd.preds.plots$Dofetilide,dd.preds.plots$Quinidine,
             dd.preds.plots$Ranolazine,dd.preds.plots$Verapamil,
             ncol=2)
```

**Figure 1**: Exposure-response plots for each treatment. Mean baseline- and placeb-corrected plasma drug concentration-dependent changes. Shaded areas show the 95% confidence intervals from model predictions. Observed data is shows as shaded dots. For clarity, observed data is grouped in 10 bins (deciles) represented by the solid circles (median concentration and mean &Delta;&Delta;ECG change), vertical error bars (95% confidence intervals of &Delta;&Delta;ECG change) and horizontal error bars (observed plasma drug concentration deciles). Horizontal lines correspond with &Delta;&Delta; = 0 ms (dashed horizontal line) and &Delta;&Delta; = 10ms (dotted horizontal line). The solid vertical line shows Cmax (vales reported in the table in next section).



#### Cmax and corresponding &Delta;&Delta; changes per treatment

Table with Cmax values for each treatment (solid vertical line in the exposure-response plots above) and corresponding predicted &Delta;&Delta; changes (mean and [95% confidence intervals])

```{r printcmaxtable, results='asis'}
Cmaxtable <- Cmaxes %>% mutate(Cmax=sprintf('%.1f %s',round(Cmax,2),PCSTRESU)) %>% select(-PCSTRESU)
ddchgtable <- ddchgatcmax %>% mutate(DDCHG=sprintf('%.1f [%.1f to %.1f] ms',round(lsmean,1),round(lower.CL,1),round(upper.CL,1))) 
Cmaxtable <- merge (Cmaxtable,ddchgtable %>% select(EXTRT,DDCHG))
## Print Cmax table
pandoc.table(Cmaxtable)
```

## JAHA manuscript Figure 3 panel K

This plot is similar to that in Figure 3.k in Vicente et al. JAHA 2015 (https://doi.org/10.1161/JAHA.114.001615).The plot shows all panels above together. This is done by rescaling the corresponding X axis to % of Cmax for each treatment (see code below and Vicente et al. for more information.)

```{r jahalikefigure, results='asis'}
## Normalize concentrations to population's Cmax
## Merge Cmaxes to the predicted double deltas data.frame and compute normalized concentration to Cmax (CONCnorm)
jaha_ddchgpreds <- merge(conc.preds.dd,Cmaxes %>% select(-PCSTRESU))  %>% group_by(EXTRT) %>% mutate(CONCnorm=PCSTRESN/max(Cmax,na.mr=T)*100)

## Generate exposure-response plot
jaha.plot.ddCHG <-ggplot()+
  geom_ribbon(data=jaha_ddchgpreds, aes(x=CONCnorm, ymin=lower.CL, ymax=upper.CL,fill=EXTRT), alpha=0.4)+ # Print model's prediction CIs
  geom_line(data=jaha_ddchgpreds,  aes(x=CONCnorm, y=lsmean, col=EXTRT, linetype=EXTRT), size=1)+ # Print model's predicted mean
  geom_hline(aes(yintercept=10), linetype=3)+ # Horizontal doted line at 10 ms
  geom_hline(aes(yintercept=0), linetype=2)+ # Horizontal dashed line at 0 ms
  geom_vline(aes(xintercept=100), linetype=1)+ # Vertical line at Cmax (i.e. 100%)
  ylab(expression(paste(Delta,Delta," \u00B1  95% CI (ms)")))+ # Y axis label
  xlab ('% of population\'s Cmax concentration')+ # X axis label
  scale_color_manual(values=treatmentcolors) + scale_fill_manual(values=treatmentcolors) + # Set treatment colors
  theme_bw() +
  theme(legend.position="bottom",legend.title=element_blank())

## Print plot
print(jaha.plot.ddCHG + scale_y_continuous(limits=yl,oob=scales::rescale_none))

figurename <- paste0('jaha-',participant,'.png')
p <- png(figurename,width=250,height=250)
print(jaha.plot.ddCHG + scale_y_continuous(limits=yl,oob=scales::rescale_none))
p<-dev.off()
```

**Figure 2**: Zoomed-in version of Figure 3.K in Vicente et al JAHA 2015.

# Appendix: R session technical information

```{r techinfo}
sessionInfo()
```